A numerical study of the solution of X-Mode equations around the hybrid resonance

### 2015-04-22 report_clean/numerics.tex
- p11 Figure 4
  dependence
much more → many more. Pike → peak. resonant → resonance.

Let us check numerically whether this holds true. 
To do so, we conduct the  numerical experiment with parameters for the problem as given in Table \ref{tab:parameters}.
\begin{table}[htb!]
\begin{tabular}{l|l}
Parameter & Value \\
\hline
$\alpha(x)$ & $\left\{\begin{array}{lr}
10, & x\leq -10,\\
-x, & -10<x\leq 5,\\
-5, & x>5.
\end{array}\right.$ \\
$\delta(x)$ & 
$\left\{\begin{array}{lr}
0, & x\leq -10,\\
4/30x+4/3,& -10<x\leq 5,\\
2, & x>5.
\end{array}\right.$ \\
$g^{inc}(-L)$ & $-2 \sqrt{2}i\exp(-22\sqrt{2}i)$\\
$\lambda$ & 
$\sqrt{10}$\\
$L$& 15\\
$H$ & 10 \\
\end{tabular}
\caption{The parameters for the problem with the resonance.}
\label{tab:parameters}
\end{table}
Denoting by $E_{x}^{h}$ the solution $E_{x}(x)$ computed on the mesh with a width $h$, and by $E_{x}^{c}$ the solution computed on a mesh multiple times finer, let 
\begin{align}
\label{eq:def_epsilon}
h_{\epsilon}=\sup\{h: \|E_{x}^{h'}-E_{x}^{c}\|<\epsilon \text{ for all } h'<h\}.
\end{align}
The computed dependence of $h_{\epsilon}$ on $\nu$ is shown in Fig.~\ref{fig:dependence}.
\begin{figure}
\begin{tabular}{cc}
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/h_nu.pdf}
&
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/res_sol.pdf}
\end{tabular}
\caption{In the left figure the depenence of $h_{\epsilon}$ as defined by (\ref{eq:def_epsilon}) on $\nu$ is demonstrated. 
In the right figure we show the computed solutions for the problem with parameters in Table \ref{tab:parameters} for different values of $\nu$. }
\label{fig:dependence}
\end{figure}
We can see that the estimate (\ref{eq:estimate_h}) is pessimistic compared to the one suggested by (\ref{eq:def_epsilon}), 
at least for given values of $\epsilon$ and for a chosen range of $\nu>0$.
\begin{remark}
It can be shown that $\|E^{\nu}_{x}\|_{L_{2}(\Omega)}\leq \frac{C}{\sqrt{\nu}},\; C>0$, 
and thus the relative error control
\bealn
 \frac{\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}}{\|E^{\nu}_{x}\|_{L_{2}(\Omega)}}\leq \epsilon
\eealn
 is ensured by choosing $h$ as $\beta_{\epsilon}\nu^{\frac{3}{4}}$.
\end{remark}

@@

We show the results of very basic simulations that the numerical strategy 
based on co-localised semi-lagrangian discretization is valid even for the resonant 
case.
The numerical tests have been kept to a minimum since much more are needed to fully validate
the concept.

The setting of the numerical results of figure \ref{fig:vasl}
is the following. We use a 7th order semi-lagrangian scheme, and the CFL is $\nu=0.5$.
The Pike is the hybrid resonance, visible on $E_x$ and $u_x$.
The total energy $\mathcal E(t)$ is the physical energy. 
The comparison with figures \ref{fig:dependence} and \ref{fig:resonance_nus_ey_x} show that the 
singular nature of the resonant is captured by this new scheme without any doubt.

>>>>

Let us check numerically whether this holds true. 
To do so, we conduct the  numerical experiment with parameters for the problem as given in Table \ref{tab:parameters}.
\begin{table}[htb!]
\begin{tabular}{l|l}
Parameter & Value \\
\hline
$\alpha(x)$ & $\left\{\begin{array}{lr}
10, & x\leq -10,\\
-x, & -10<x\leq 5,\\
-5, & x>5.
\end{array}\right.$ \\
$\delta(x)$ & 
$\left\{\begin{array}{lr}
0, & x\leq -10,\\
4/30x+4/3,& -10<x\leq 5,\\
2, & x>5.
\end{array}\right.$ \\
$g^{inc}(-L)$ & $-2 \sqrt{2}i\exp(-22\sqrt{2}i)$\\
$\lambda$ & 
$\sqrt{10}$\\
$L$& 15\\
$H$ & 10 \\
\end{tabular}
\caption{The parameters for the problem with the resonance.}
\label{tab:parameters}
\end{table}
Denoting by $E_{x}^{h}$ the solution $E_{x}(x)$ computed on the mesh with a width $h$, and by $E_{x}^{c}$ the solution computed on a mesh multiple times finer, let 
\begin{align}
\label{eq:def_epsilon}
h_{\epsilon}=\sup\{h: \|E_{x}^{h'}-E_{x}^{c}\|<\epsilon \text{ for all } h'<h\}.
\end{align}
The computed dependence of $h_{\epsilon}$ on $\nu$ is shown in Fig.~\ref{fig:dependence}.
\begin{figure}
\begin{tabular}{cc}
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/h_nu.pdf}
&
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/res_sol.pdf}
\end{tabular}
\caption{In the left figure the dependence of $h_{\epsilon}$ as defined by (\ref{eq:def_epsilon}) on $\nu$ is demonstrated. 
In the right figure we show the computed solutions for the problem with parameters in Table \ref{tab:parameters} for different values of $\nu$. }
\label{fig:dependence}
\end{figure}
We can see that the estimate (\ref{eq:estimate_h}) is pessimistic compared to the one suggested by (\ref{eq:def_epsilon}), 
at least for given values of $\epsilon$ and for a chosen range of $\nu>0$.
\begin{remark}
It can be shown that $\|E^{\nu}_{x}\|_{L_{2}(\Omega)}\leq \frac{C}{\sqrt{\nu}},\; C>0$, 
and thus the relative error control
\bealn
 \frac{\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}}{\|E^{\nu}_{x}\|_{L_{2}(\Omega)}}\leq \epsilon
\eealn
 is ensured by choosing $h$ as $\beta_{\epsilon}\nu^{\frac{3}{4}}$.
\end{remark}

@@

We show the results of very basic simulations that the numerical strategy 
based on co-localised semi-lagrangian discretization is valid even for the resonant 
case.
The numerical tests have been kept to a minimum since many more are needed to fully validate
the concept.

The setting of the numerical results of figure \ref{fig:vasl}
is the following. We use a 7th order semi-lagrangian scheme, and the CFL is $\nu=0.5$.
The peak is the hybrid resonance, visible on $E_x$ and $u_x$.
The total energy $\mathcal E(t)$ is the physical energy. 
The comparison with figures \ref{fig:dependence} and \ref{fig:resonance_nus_ey_x} show that the 
singular nature of the resonance is captured by this new scheme without any doubt.
